Study of Intra- and Inter-Job Interference with Different Job allocations

### 2015-10-14 submission.tex
changes based on Lan's comments

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Study of Intra- and Inter-Job Interference for Improving Allocation on HPC Systems}
% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{

@@

This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resource. The batch scheduler is responsible for assigning job the amount of nodes it requires. On the widely used torus-connected HPC systems \cite{bgq}\cite{tofu}\cite{titan}, two allocation strategies are commonly used. In the partition based allocation, used in Blue Gene series systems, he scheduler assigns each user job a compact and contiguous set of computing nodes. This strategy favors application performance by providing application with isolated partitions and exclusive network connections, thus reducing network contention caused by concurrently running jobs. However, this strategy can cause internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization that is otherwise possible. The other strategy is non-contiguous allocation, used by Cray XT/XE series, in which free nodes are assigned to user jobs whether they are contiguous or not. Non-contiguous allocation eliminates internal and external fragmentation as seen in partition-based systems, thereby leading to high system utilization. However, it introduces other problems, for example scattering application processes all over the system, causing both inter-job and intra-job contention. The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive jobs.

@@

%=========come back and work from here=============
Rather than choosing blindly between partition-based or non-contiguous allocation, the new allocation strategy choses resources based on a job's communication pattern. With detail analysis of application's communication patterns, our proposed allocation strategy obtains a clear view as to which processes are tightly connected and conduct intensive communication. The ``neighborhood communication" consists of such tightly coupled processes should get a compact allocation. Then, the allocation strategy only needs to guarantee the locality of the subset of the nodes where those ``neighborhood communication" would reside.

In this work, we study the impact of different allocation strategies on intra/inter-job interference on Torus network. Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance. On the recent Top500 list, 6 of the top 10 supercomputers use a high-radix torus-interconnected network \cite{top500}. The current generation of IBM Blue Gene/Q (BG/Q) supercomputer, such as Mira at Argonne Nation Laboratory and Sequoia at  Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}. The K computer from Japan uses the “Tofu” system, which is a 6D mesh/torus topology \cite{tofu}. Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}. Although the analysis are based on torus network, the idea conveyed from this work is applicable to networks with different topologies.

We selected three signature applications from DOE Design Forward Project\cite{design forward webpage} as examples to conduct detailed study about their communication patterns. We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011}, from Argonne National Laboratory as a research vehicle to evaluate the performance of our communication pattern aware allocation policy on torus networks. We analyze each application's performance running on torus networks with different dimensionality and interconnected bandwidth. Then, we extend CODES with a Job Mapping API that supports different allocation policies on torus network. Different allocation strategies on torus network are studied with the help of this new Job Mapping API. We analyze the intra/inter-job interference by simulating those three applications running exclusively or concurrently on torus network with different allocation strategies. According to results of our comprehensive exploration, making allocation with consideration about job's communication pattern is a very promising allocation strategy. We believe the idea presented in this work would be the guideline for the design of future HPC batch job scheduler and resource management module.

@@

A parallel application usually conforms to a combination of several basic communication patterns \cite{hpdc2015-ornl}. At its different execution phases, the application's communication behavior may follow different basic patterns respectively. Collective communication is often used for data distribution and collection among processes. Nearest Neighbor communication is commonly seen in lots of structured and unstructured mesh physics packages. There are many profiling tools available to capture communication patterns from parallel applications, such as Tuning and Analysis Utilities(TAU) \cite{tau}, mpiP \cite{mpip}. They can help analysis parallel application, providing information like the percentage of different MPI operations of the applications, communication topology, the amount data transferred between processes.

In this work, we provide detailed analysis about the communication pattern of three DOE MiniApps. MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}. Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures. We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter. Their representative communication patterns that are commonly seen in the HPC system workload. The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{design forward webpage}.

@@

AMG is the MiniApp of Algebraic Multigrid Solver, which is a parallel algebraic multigrid solver for linear systems arising from problems on unstructured mesh physics packages. It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}. The dominant communication pattern is this regional communication with decreasing message size for different parts of the multigrid v-cycle.

Figure \ref{fig: amg communication topology} shows the communication topology of a small scale AMG with 27 MPI Ranks. Here we prefer to show the communication topology with small scale version of the MiniApp because the dominant communication pattern of the application doesn't change with its scale. We can make the observation from the upper-left of the figure that each rank in AMG has intense communication between 3 other ranks, such as rank 0 between rank 1, 3 and 9, rank 1 between 2, 4 and 10, etc. And this relation is also symmetric along the main diagnal. The dominant communication pattern is quite obvious when we identify this relation between ranks.

Figure \ref{fig: amg data trans} shows the amount of data each rank transferred in AMG. The blue line shows that total data amount transferred from each rank, while the red shows the amount of received data and green shows the amount of send data. The first thing we can spot from the figure is that for each rank, the amount of data sent and received are basically the same, which also well explain the symmetry of the application's dominant communication pattern. And the second thing we can find about AMG is that the data transfer amount also has a symmetric pattern. Rank 0 has the same amount of data transfer as Rank 26, Rank 1 as of Rank 25, etc.

@@

The second MiniApp we have is Crystal Router, which is the extracted communication kernel of the full application Nek5000. Nek5000 \cite{crystalrouter} is a spectral element CFD application developed at Argonne National Laboratory. It features spectral element multigrid solvers coupled with a highly scalable, parallel coarse-grid solver that widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos. The communication kernel of Nek5000, CrystalRouter demonstrates the "many-to-many" communication pattern through scalable multi-stage communication process. \textcolor{yellow}{The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.}

As we can see from Figure \ref{fig: cr communication topology}, every 5-rank is a compact community, where data transfer is very intensive. Beside the "5-rank community", there is "pair-wise" global data transfer in CrystalRouter. Obviously, there is data transfer alone the diagnal. \textcolor{yellow}{This is because messages would be buffered on their node of origin until a synchronization point was reached when a single system call sent every message to its destination in one pass.} When the scale of Crystal Router increase to 100, the "5-rank community" is still the dominant pattern in a hierarchical manner. The same pattern can be observed from the global data transfer. Our observation about the dominant communication pattern of CrystalRouter is the multi-stage "5-rank community" and "pair-wise" global data transfer happen hierarchically as the scale of the application increase.

Figure \ref{fig: cr data trans} shows the amount of data each rank transferred in Crystal Router. The locality pattern of "5-rank community" can still be spotted in this figure. The spiky shape occurs in the frequency of every 5 rank. Unlike AMG, the there is variance between the amount of data send and received for each rank. This is due to the specific feature of Crystal Router's communication pattern.

@@

\subsection{MultiGrid}
\label{sec:multigrid}
Different from AMG, MultiGrid is geometric multigrid v-cycle from production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes. MultiGrid conforms to Many-to-Many communication pattern with decreasing message size and collectives for different parts of the multigrid v-cycle. It is widely used for structured grid physics packages. Figure \ref{fig: mg communication topology} shows the communication topology of MultiGrid and Figure\ref{fig: mg data trans} shows the the amount of data transferred between ranks in MultiGrid.

@@

MultiGrid represents a group of applications whose dominant communication pattern is Many-to-Many (some literatures also refer it as All-to-All \cite{hpdc2015-ornl}).

@@

It is impossible to accurately study the interference between concurrently running jobs with different allocation strategies on production systems. There are a couple of reasons for that. First, the allocation strategy used on production machine is part of the system software, which can not be changed by user. Even the system administrator is not authorized to make that change. Second, it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference then compare the results to that in the presence of interference. Due to those reasons, we resort to simulation for this work.

There are couple of simulation platform to study the behavior of large scale systems. SimGrid from Inria is a versatile simulation tool sets that provide simulation for Grids, Clouds and HPC systems \cite{simgrid}. The Structural Simulation Toolkit (SST) from Sandia National Lab was developed to explore innovations in highly concurrent systems where the ISA, microarchitecture, and memory interact with the programming model and communications system\cite{sst}.

A simulation toolkit named CODES (Co-Design of Multi-layer Exascale Storage Architecture) from Argonne National Lab enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}. CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}. CODES support both torus and dragonfly network with high fidelity flit-level simulation. In this work, we only use torus network to show the impact of node allocation to job with different dominant communication patterns. CODES has this network workload component that capable of conducting trace-driven simulation. It can take real MPI application trace generated by SST DUMPI \cite{sst} to drive CODES network models. We implemented a Job Mapping API for CODES, which can provide flexible job allocation strategies. The new API can help us to explore the impact of different allocation strategies to those three applications in Section \ref{sec:application study}

\section{Experiment}
\label{sec:experiment results}

CODES can provide Torus network model with good scalability and high fidelity \cite{codes-torus}. In this section, we will first study the communication behavior of three applications on torus network with different dimensionality and bandwidth configurations. Then, we study the intra- and inter-job interference with these three applications by running them exclusively and concurrently with different allocation strategies.

\subsection{Configuration Study }
\label{sec:config study}

The topology of torus network is k-ary n-cube, with $k^n$ nodes in total arranged in an n-dimensional grid having k nodes in each dimension. Each node has 2$\times$ n direct linked neighbor nodes. The torus network performance is determined by its dimensionality and link bandwidth. As the increase of the dimensionality of torus network, so does the number of links connected with each node. The increased aggregated bandwidth of each node will definitely reduce the data transfer time of each rank in the application. Figure \ref{fig:dimensionality-study} shows the performance of the AMG, CrystalRouter and MultiGrid on a 2K node torus network model with a 3D torus (16x16x8), a 5D torus (8x4x4x4x4), and a 7D torus (4x4x4x4x2x2x2). The bandwidth of direct link between nodes is 2GiB/s in each direction, thus, the aggregated bandwidth is 12GiB/s per node in 3D torus, 20GiB/s per node in 5D torus, and 28GiB/s per node in 7D torus.

We can see from Figure \ref{fig:dimensionality-study} that the data transfer time of those three applications are greatly reduced as the dimensionality increases. The increased aggregated bandwidth of each node can accelerate the transfer of data. This can justify the fidelity and consistency of the torus network model. We allocate each rank within each application linearly on the 3D torus. However, due to the variance of data transfer volume of each rank, AMG's data transfer time has great dispersion, shown as the outliers above the boxes in Figure \ref{fig:dimen-amg}. There is no such great dispersion for CrystalRouter and MultiGrid, shown in Figure \ref{fig:dimen-cr} and \ref{fig:dimen-mg}, since no great variance of data transfer amount of each rank in these applications.

@@

\subsection{Interference Analysis}
\label{sec:interference}

@@

\textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources. Such interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}. Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than partition-based system. Application's communication time can vary from 36\% faster to 69\% slower due to inter-job interference when running on non-contiguous allocation \cite{abhinav-sc13}.

@@

We can get observations that 
\begin{enumerate}
    \item The compact allocation can not guarantee the best performance for every application. 
    \item The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefer compact allocation. 
    \item The allocation unit size should be decided according to application's communication pattern. Unit size that big enough to preserve the locality of the application will result better performance. 
\end{enumerate}

@@

\subsubsection{Intra-Job Interference Analysis}
\label{sec: introjob}

@@

\subsubsection{Inter-Job Interference Analysis}
\label{sec: interjob interference study}

@@

The data transfer time of AMG in Figure \ref{fig:interjob-amg-box} keeps stable between allocation unit size of 16(bit) and 8(medium), since the locality of AMG is 6-rank based. When the unit size reduce to 2(small), AMG suffers prolong data transfer time by about 10\%. CrystalRouter is more sensitive to allocation unit size. The best unit size would be the one big enough to accommodate its 5-rank locality. Figure \ref{fig:interjob-cr-box} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16. When the unit size reduce to 2-node, the communication become less efficiency and take 15\% more times for transferring data.

@@

The most promising way to utilize such knowledge is to make smart allocation based on job's dominant communication pattern. Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy. Whether it use partition-based or non-contiguous policy, means it chooses the benefit of either system or job over the other. Some propose compromised policies \cite{lo}\cite{pascual}\cite{yang-cluster14} to use fix allocation unit size and  make the allocation for job blindly without knowing job's communication information.

Based on the findings in this work, we propose the idea of smart job allocation with consideration about job's communication pattern. As the dominant communication pattern of parallel applications on HPC system can be summarized into a couple of categories, the deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern. When an application with dominant communication that is intensive all-to-all submitted to the system, it should be granted with compact allocation with exclusive network provision, so that it won't suffer performance loss due to interference. If job with dominant communication pattern like Nearest Neighbor or others with intensive small scale communication, rather than assign it with unnecessary compact allocation and exclusive network provision, non-contiguous allocation with proper unit size would serve it equally well. No degradation will be introduced to job's performance and system utilization will be improved.

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Study of Intra- and Inter-Job Interference with Different Job allocations}
% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{

@@

This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resource. The batch scheduler is responsible for assigning job the amount of nodes it requires. On the widely used torus-connected HPC systems \cite{bgq}\cite{tofu}\cite{titan}, two allocation strategies are commonly used. In the partition based allocation, used in Blue Gene series systems, he scheduler assigns each user job a compact and contiguous set of computing nodes. This strategy favors application performance by providing application with isolated partitions and exclusive network connections, thus reducing network contention caused by concurrently running jobs. However, this strategy can cause internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization that is otherwise possible. The other strategy is non-contiguous allocation, used by Cray XT/XE series, in which free nodes are assigned to user jobs whether they are contiguous or not. Non-contiguous allocation eliminates internal and external fragmentation as seen in partition-based systems, thereby leading to high system utilization. However, it introduces other problems, for example scattering application processes all over the system, causing both inter-job and intra-job contention. The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones.

@@

We envision that future HPC will adopt a flexible job allocation mechanism which combines the best of both partition-based allocation and non-contiguous allocation. Such flexible mechanism would make allocation for jobs based on their communication patterns. With detail analysis of application's communication patterns, it can be identified which processes are tightly connected and conduct intensive communication within each job. Rather than assign each job a compact partition-based allocation or non-contiguous allocation, the ``neighborhood communication" consists of such tightly coupled processes should get a compact allocation. The new allocation mechanism only needs to guarantee the locality of the subset of the nodes where those ``neighborhood communication" would reside.

In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems. Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance. On the recent Top500 list, 6 of the top 10 supercomputers use a high-radix torus-interconnected network \cite{top500}. The current generation of IBM Blue Gene/Q (BG/Q) supercomputer, such as Mira at Argonne Nation Laboratory and Sequoia at  Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}. The K computer from Japan uses the “Tofu” system, which is a 6D mesh/torus topology \cite{tofu}. Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}. Although the analysis are based on torus network, the idea conveyed from this work is applicable to networks with different topologies.

We selected three signature applications from DOE Design Forward Project\cite{design forward webpage} as examples to conduct detailed study about their communication patterns. We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011}, from Argonne National Laboratory as a research vehicle to evaluate the performance of these applications with various allocations. We analyze the intra- and inter-job interference by simulating applications running exclusively or concurrently with different allocations.

\textbf{replace this by summarizing a couple of key findings!!} According to results of our comprehensive exploration, making allocation with consideration about job's communication pattern is a very promising allocation strategy.

@@

A parallel application usually conforms to a combination of several basic communication patterns \cite{hpdc2015-ornl}. At its different execution phases, the application's communication behavior may follow different basic patterns respectively. There are many profiling tools \cite{tau}\cite{mpip}\cite{scala}\cite{sst}\cite{oxbow} available to capture information regarding communication patterns of parallel applications. They can help analyze parallel applications' communication behavior, providing information like the percentage of different MPI operations of the applications, communication topology, the amount data transferred between processes.

There are many applications running on HPC systems like Mira \cite{bgq} and Titan \cite{titan}. Most of these applications are parallel and communication intensive, and conform to different communication patterns. In this work, we select three representative applications from DOE Design Forward project \cite{design forward webpage}. Each application conforms to a distinctive communication pattern that is commonly seen from applications in HPC workload. We believe these three applications can represent a big group of applications that conform to similar communication patterns. The applications studied in this work are Algebraic MultiGrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter. The communication pattern figures we present in Section \ref{sec:amg}, \ref{sec:crystalrouter}, \ref{sec:multigrid} are generated with the IPM \cite{ipm} data from \cite{design forward webpage}.

% MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}. Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures. We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter. Their representative communication patterns that are commonly seen in the HPC system workload. The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{design forward webpage}.

@@

AMG is the MiniApp of Algebraic MultiGrid Solver, which is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages. It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}. The dominant communication pattern is this regional communication with decreasing message size for different parts of the multi-grid v-cycle.

Figure \ref{fig: amg communication topology} shows the communication topology of a small scale AMG with 216 MPI Ranks. Here we show the communication topology with small scale version because the dominant communication pattern of the application doesn't change with scale. We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with with its 6 neighbors unless it is on the boundary. The dominant communication pattern is quite obvious when we identify this relation between ranks.

Figure \ref{fig: amg data trans} shows the amount of data each rank transferred in AMG. The red line shows the amount of data sent by each rank, the blue line shows the amount of received data. The data transferring of each rank is between 7MB to 16MB.

@@

The second MiniApp we have is CrystalRouter, the extracted communication kernel of the full application Nek5000 \cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory. It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos. CrystalRouter demonstrates the "many-to-many" communication pattern through scalable multi-stage communication process. The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.

The collective communication in CrystalRouter utilizing a recursive doubling approach. Ranks in CrystalRouter conform to a n-dimensional hypercube and recursively splitting into (n-1)-dimensional hypercubes, with communication happening with respect to the links that are severed. The pattern of this recursive communication can be found in Figure \ref{fig: cr communication topology}. By nature of the logarithmic distribution, more data transferring will be conducted within the local neighborhood, which consists of 5 ranks.

Figure \ref{fig: cr data trans} shows the amount of data each rank transferred in CrystalRouter. The spiky shape indicates there is intensive data transferring in the local neighborhood. There is great variance of data transferring among the ranks in CrystalRouter, which is between 20MB to 36MB. This is due to the specific feature of CrystalRouter's communication pattern.

@@

\subsection{MultiGrid}
\label{sec:multigrid}

@@

MultiGrid is geometric multi-grid v-cycle from production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes. MultiGrid conforms to Many-to-Many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle. It is widely used for structured grid physics packages.

Figure \ref{fig: mg communication topology} shows the communication topology of MultiGrid with 125 ranks. We can see intensive communication along the diagnal that conforms to 3D Nearest Neighbor communication pattern. However, there are also considerable amount of data transferring on both sides of the diagnal, which means the dominant communication pattern of MultiGrid is ``Many-to-Many".

Figure\ref{fig: mg data trans} shows the the amount of data transferred between ranks in MultiGrid. The data transferring of each rank is relatively small and stable, which is between 2.8MB to 2.9MB.

MultiGrid represents a group of applications whose dominant communication pattern is Many-to-Many (some literatures also refer it as All-to-All \cite{hpdc2015-ornl}).  Applications with such similar dominant communication pattern are like FillBoundary and MiniDFT \cite{design forward webpage}.

@@

It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context. There are a couple of reasons for that. First, the allocation strategy used on production machine is part of the system software, which can not be changed by user. Even the system administrator is not authorized to make that change. Second, it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference then compare the results to that in the presence of interference. Due to those reasons, we resort to simulation for this work.

%There are couple of simulation platform to study the behavior of large scale systems. SimGrid from Inria is a versatile simulation tool sets that provide simulation for Grids, Clouds and HPC systems \cite{simgrid}. The Structural Simulation Toolkit (SST) from Sandia National Lab was developed to explore innovations in highly concurrent systems where the ISA, microarchitecture, and memory interact with the programming model and communications system\cite{sst}.

A simulation toolkit named CODES (Co-Design of Multi-layer Exascale Storage Architecture) from Argonne National Lab enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}. CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}. CODES support both torus and dragonfly network with high fidelity flit-level simulation. CODES has this network workload component that capable of conducting trace-driven simulation. It can take real MPI application trace generated by SST DUMPI \cite{sst} to drive CODES network models. In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems. Since torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
%We implemented a Job Mapping API for CODES, which can provide flexible job allocation strategies. The new API can help us to explore the impact of different allocation strategies to those three applications in Section \ref{sec:application study}

\section{Configuration Study }
\label{sec:config study}

In this section, we study the communication behavior of three applications on torus network with different dimensionality and bandwidth configurations.

The topology of torus network is k-ary n-cube, with $k^n$ nodes in total arranged in an n-dimensional grid having k nodes in each dimension. Each node has 2$\times$ n direct linked neighbor nodes. The torus network performance is determined by its dimensionality and link bandwidth. As the increase of the dimensionality of torus network, so does the number of links connected with each node. The increased aggregated bandwidth of each node will definitely reduce the data transfer time of each rank in the application. Figure \ref{fig:dimensionality-study} shows the performance of the AMG, CrystalRouter and MultiGrid on a 2K node torus network model with a 3D torus (16x16x8), a 5D torus (8x4x4x4x4), and a 7D torus (4x4x4x4x2x2x2). The bandwidth of direct link between nodes is 2GiB/s in each direction \cite{bgq}, thus, the aggregated bandwidth is 12GiB/s per node in 3D torus, 20GiB/s per node in 5D torus, and 28GiB/s per node in 7D torus.

We can see from Figure \ref{fig:dimensionality-study} that the data transfer time of those three applications are greatly reduced as the dimensionality increases. The increased aggregated bandwidth of each node can accelerate the data transfer. We allocate each rank within each application linearly on the 3D torus. However, due to the variance of data transfer volume of each rank, AMG's data transfer time has great dispersion, shown as the outliers above the boxes in Figure \ref{fig:dimen-amg}. There is no such great dispersion of MultiGrid, shown in Figure \ref{fig:dimen-mg}, since there is no great variance of data transfer of each rank. The data transfer of each rank in MultiGrid varies slightly  between 2.8MB to 2.9MB. \textcolor{red}{explain crystalrouter}.

@@

\section{Interference Analysis}
\label{sec:interference}

@@

\textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources. Such interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}. Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than partition-based system. Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to inter-job interference when running on non-contiguous allocation \cite{abhinav-sc13}.

@@

We can get observations that 
\begin{enumerate}
    \item The compact allocation may not guarantee the best performance for every application. 
    \item The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefer compact allocation. 
    \item The allocation unit size should be decided according to application's communication pattern. Unit size that big enough to preserve the locality of the application will result better performance. 
\end{enumerate}

@@

\subsection{Intra-Job Interference Analysis}
\label{sec: introjob}

@@

\subsection{Inter-Job Interference Analysis}
\label{sec: interjob interference study}

@@

The data transfer time of AMG in Figure \ref{fig:interjob-amg-box} keeps stable between allocation unit size of 16 (big) and 8 (medium), since the locality of AMG is 6-rank based. When the unit size reduce to 2 (small), AMG suffers prolong data transfer time by about 10\%. CrystalRouter is more sensitive to allocation unit size. The best unit size would be the one big enough to accommodate its 5-rank locality. Figure \ref{fig:interjob-cr-box} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16. When the unit size reduce to 2-node, the communication become less efficiency and take 15\% more times for transferring data.

@@

The most promising way to utilize such knowledge is to make smart allocation based on job's dominant communication pattern. Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy. Whether it uses partition-based or non-contiguous policy, it chooses the benefit of either system or job over the other. Some proposed compromised policies \cite{lo}\cite{pascual}\cite{yang-cluster14} to use fix allocation unit size and  make the allocation for job blindly without knowing job's communication information.

providing insights for ....
Based on the findings in this work, we propose the idea of smart job allocation with consideration about job's communication pattern. As the dominant communication pattern of parallel applications on HPC system can be summarized into a couple of categories, the deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern. When an application with dominant communication that is intensive all-to-all submitted to the system, it should be granted with compact allocation with exclusive network provision, so that it won't suffer performance loss due to interference. If job with dominant communication pattern like Nearest Neighbor or others with intensive small scale communication, rather than assign it with unnecessary compact allocation and exclusive network provision, non-contiguous allocation with proper unit size would serve it equally well. No degradation will be introduced to job's performance and system utilization will be improved.
